### 2014-09-22 ApTranscriptome_TR.Rmd

pandoc.table(sigtable, style="rmarkdown", caption = "Number of transcripts with expression that depends on species, temperature or their interaction at 5% FDR  out of 98,186 total transcripts.")
```

@@

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'Ar.type'], Ap.response.type[ , 'A22.type'])
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

@@

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, qlslope, qlmean, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.BP.resultParentChild, resp.stress.responsive.count, t.mean, t.slope, t.curvature, t.wiggle, t.varint, t.varbim, T_on.high.ttest, T_on.low.ttest, Ap.gsea.union, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, file = "ms.RData")
```

>>>>

pandoc.table(sigtable, style="rmarkdown", big.mark = ",", caption = "Number of transcripts with expression that depends on species, temperature or their interaction at 5% FDR  out of 98,186 total transcripts.")
```

@@

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Acar = Ap.response.type[ , 'Ar.type'], Apic = Ap.response.type[ , 'A22.type'])
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

@@

The question of biological interest is whether the marginal frequencies differ between the two species. Statistically, this can be addressed using the generalized [McNemar's test](http://en.wikipedia.org/wiki/McNemar's_test) of marginal homogeneity.

```{r mh_test}
mh.test <- mh_test(tt2)

# calculate contribution of each off-diagonal to Z~0~
d1 <- sum(tt2[1,2:5]) - sum(tt2[2:5,1])
d2 <- sum(tt2[2,c(1,3:5)]) - sum(tt2[c(1,3:5),2])
d3 <- sum(tt2[3,c(1:2,4:5)]) - sum(tt2[c(1:2,4:5),3])
d4 <- sum(tt2[4,c(1:3,5)]) - sum(tt2[c(1:3,5),4])
d5 <- sum(tt2[5,1:4]) - sum(tt2[1:4,5])

# proportion of test statistic due to off-diagonals
dsum <- sum(abs(d1), abs(d2), abs(d3), abs(d4), abs(d5))
d1/dsum
d2/dsum
d3/dsum
d4/dsum
d5/dsum

# e2 (Acar Intermediate) contributes ~44% to Z~0~
```

```{r mh_plot}
# overall mean for each class
rs <- rowSums(tt2)
cs <- colSums(tt2)
(gm <- (rs + cs) / sum(tt2 * 2))

# calculate observed values for each cell using overall mean for each expression type
Ec <- outer(gm, gm, "*") * sum(tt2)

# get deviations of observed from expected
Ec.dev <- tt2 - Ec

# calculate chi-squared deviation 
Ec.cells <- sign(tt2 - Ec) * (tt2 - Ec)^2 / Ec

md <- melt(Ec.cells)

qplot(x=Apic, y=Acar, data=md, fill=value, geom="tile", ylim = rev(levels(md$Acar))) + 
  scale_fill_gradient2(limits=c(-600, 600)) 
```

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, md, qlslope, qlmean, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.BP.resultParentChild, resp.stress.responsive.count, t.mean, t.slope, t.curvature, t.wiggle, t.varint, t.varbim, T_on.high.ttest, T_on.low.ttest, Ap.gsea.union, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, file = "ms.RData")
```
